---
title: "Analyse pilot data"
author: "Jörn Alexander Quent"
date: "31 January 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 30)
```

## Introduction
This script is about to check whether the parameters of the experiment (such as time in the VE, number of targets objects etc.) works to avoid either floor or ceiling effects. My aim is to use vectorisation more often while coding. This is what I try here. The data is not real. 

## Libraries
```{r}
library(data.table)
library(plyr)
library(ggplot2)
```

## Loading data
```{r}
# Using fread instead of read.table because it is faster but rbindlist(lapply(fileNames, read.table)) would also work. This is my attempt to use vectorisation more often. 
population        <- 1:2
fileNames         <- paste('data/retrievalTask_', as.character(population),'.dat', sep = '')
dataRetrievalTask <- rbindlist(lapply(fileNames, fread))

# Variable and objects names
names(dataRetrievalTask) <- c('subNo', 'date', 'time', 'trial','objectNumber', 'targetLocation', 'targetRankPre', 'foil1Location', 'foil1RankPre', 'foil2Location', 'foil2RankPre', 'AFCPreTime','rtAFC', 'resAFC', 'accAFC', 'conPreTime', 'rtCon', 'resCon', 'left', 'middle', 'right')
objectNames              <- c('microwave','kitchen roll','saucepan', 'toaster','bowl of fruits','tea pot','knife','mixer','bread','glass jug','mug','dishes','towels','toy','pile of books','umbrella','hat','helmet','calendar','fan')

# Assign factor labels
dataRetrievalTask$objectName <- factor(dataRetrievalTask$objectNumber, labels = objectNames)
dataRetrievalTask$left       <- factor(dataRetrievalTask$left, labels = c('target', 'foil1', 'foil2'))
dataRetrievalTask$middle     <- factor(dataRetrievalTask$middle, labels = c('target', 'foil1', 'foil2'))
dataRetrievalTask$right      <- factor(dataRetrievalTask$left, labels = c('target', 'foil1', 'foil2'))
```

## Confidence decision
```{r}
# Legend: 1 = Did not see object | 2 = Guess the object was there | 3 = Know the object was there
# For each participant
ddply(dataRetrievalTask, c('subNo'), summarise, notSeen = table(resCon)[1], guessed = table(resCon)[2], knew = table(resCon)[3])

# Overall
table(dataRetrievalTask$resCon)
```

## Accuracy for AFC decision
### Not exlcuding trials with 'didn't see the object' responses

```{r}
# For each participant
agg1 <- ddply(dataRetrievalTask, c('subNo'), summarise, accAFC = mean(accAFC))
agg1

# overall mean 
mean(dataRetrievalTask$accAFC)
# Overal median
median(dataRetrievalTask$accAFC)
# SD
sd(dataRetrievalTask$accAFC)
# Min
min(dataRetrievalTask$accAFC)
# Max
max(dataRetrievalTask$accAFC)

```

### Exlcuding trials with 'didn't see the object' responses (subset 1)
```{r}
# For each participant
dataRetrievalTask_subset1 <- subset(dataRetrievalTask, dataRetrievalTask$resCon != 3)
agg2                      <- ddply(dataRetrievalTask_subset1, c('subNo'), summarise, accAFC = mean(accAFC))
agg2

# overall mean 
mean(dataRetrievalTask_subset1$accAFC)
# Overal median
median(dataRetrievalTask_subset1$accAFC)
# SD
sd(dataRetrievalTask_subset1$accAFC)
# Min
min(dataRetrievalTask_subset1$accAFC)
# Max
max(dataRetrievalTask_subset1$accAFC)
```

### Visualisation 
```{r}
agg1$include1Responses <- rep('yes', length(population))
agg2$include1Responses <- rep('no', length(population))
aggCombined            <- rbind(agg1, agg2)
aggOverall             <- ddply(aggCombined, c('include1Responses'), summarise, mean = mean(accAFC))


ggplot(aggCombined, aes(x = include1Responses, y = accAFC)) + 
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) + geom_hline(yintercept = 1/3) + 
  geom_segment(aes(x = 0.8, y = aggOverall$mean[1], xend = 1.2, yend = aggOverall$mean[1])) +
  geom_segment(aes(x = 1.8, y = aggOverall$mean[2], xend = 2.2, yend = aggOverall$mean[2])) + 
  annotate('text', x = 0.6, y = (1/3) - 0.03, label = 'Chance (1/3)') + 
  labs(y = 'Accuracy (3AFC)', x = "Includes 'Did not see object' responses") + 
  coord_cartesian(ylim = c(0, 1)) + 
  theme(plot.margin = margin(10, 10, 10, 10))
```