---
title: "Checking slopes to test consistency across participants"
author: "Jörn Alexander Quent"
date: "16 June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap')
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.width = 10, 
                      fig.height= 7, 
                      dpi=300, 
                      out.width="1200px", 
                      out.height="700px")
options(scipen = 30)
```

# Aim
I am currently writing up my first year report (see [here](https://jaquent.me/category/project-1/) for information about this project). Currently, the data set includes 23 participants and I am really happy to report that I found a significant quadratic relationship between object/location expectancy and memory performance. Now, the question is whether the quadratic effect is consistent across participants or whether some participants even show the reverse effect. 

# Libraries & data
```{r}
library(ggplot2)
library(lmerTest)

# Loading 
load("data/exp2Data.RData")
dataSchemaVR2 <- combData
rm(combData)

# Preparation
dataSchemaVR2                           <- subset(dataSchemaVR2, subNum >= 25)
dataSchemaVR2_Scaled                    <- dataSchemaVR2
dataSchemaVR2_Scaled$objLocTargetRating <- scale(dataSchemaVR2_Scaled$objLocTargetRating)
dataSchemaVR2_AFC                       <- subset(dataSchemaVR2_Scaled, dataSchemaVR2_Scaled$resCon != 0)
```


# Model
In order to test the consistency of the quadratic effect I allowed the slopes to vary across participants. The model is a logistic regression/Generalized Linear Mixed-Effects model with a random intercept and random linear and quadratic slope predicting accuracy on each trial for each participant.  

```{r}
schemaVR2_model15_slopes <- glmer(accAFC ~  objLocTargetRating +
                                    I(objLocTargetRating*objLocTargetRating) +
                                    (objLocTargetRating | subNum) + 
                                    (I(objLocTargetRating*objLocTargetRating) | subNum),
                                  data = dataSchemaVR2_AFC,
                                  family = binomial,
                                  control = glmerControl(optimizer = "bobyqa"),
                                  nAGQ = 1)

summary(schemaVR2_model15_slopes)
coef(schemaVR2_model15_slopes)$subNum
slopeMean <- round(mean(unlist(coef(schemaVR2_model15_slopes)$subNum[3])), 2)
slopeSD   <- round(sd(unlist(coef(schemaVR2_model15_slopes)$subNum[3])), 2)
```

The mean quadratic slope is `r slopeMean` and the standard deviation is only `r slopeSD`, which shows that the effect is remarkably consistent across participants. There are no participants, for whom the effect is reversed. 

# Visualisation
The slopes are plotted below to visualise them. 

```{r} 
minExp <- min(dataSchemaVR2_AFC$objLocTargetRating)
maxEp  <- max(dataSchemaVR2_AFC$objLocTargetRating)
n      <- dim(coef(schemaVR2_model15_slopes)$subNum)[1]

x      <- seq(minExp, maxEp, 0.01)
nCol   <- length(x)
x      <- matrix(rep(x, n), nrow = n, ncol = nCol)

# Important: The coefficients need to be unlisted otherwise the multiplication doesn't work. 
b0     <- unlist(coef(schemaVR2_model15_slopes)$subNum[1])
b1     <- unlist(coef(schemaVR2_model15_slopes)$subNum[2])
b2     <- unlist(coef(schemaVR2_model15_slopes)$subNum[3])

y      <- b2*(x*x) + b1*x + b0

slopes        <- data.frame(subNum = rep(1:n, each = nCol), x = c(t(x)), y = c(t(y)))
slopes$subNum <- as.factor(slopes$subNum)

ggplot(slopes, aes(x = x, y = y, group = subNum, colour = subNum)) + 
  geom_line(show.legend = FALSE) + 
  labs(y = 'Accuracy', x = 'Expectancy')
```